Dynamical Properties of Molecular Cloud Complexes in Galaxies at the Epoch of Reionization

### 2019-07-05 WRITEUP/ms.tex

\begin{abstract}
We study the properties of molecular cloud complexes (MCCs) at the Epoch of Reionization (EoR) using cosmological zoom-in simulations. We identify MCCs in a $z\simeq 6$ prototypical galaxy (``\flower'') using an H$_2$ density-based clump finder.
%
We compare their mass, size, velocity dispersion, gas surface density, and virial parameter ($\alpha_{\rm vir}$) to \obs. In \flower, the typical MCC mass and size are $M_{\rm gas} \simeq 10^{6.5}$\,\Msun and $R\simeq45-100$\,pc, which are comparable to those found in nearby spirals and starburst galaxies. MCCs are highly supersonic and supported by turbulence, with r.m.s velocity dispersions of $\sigma_{\rm gas}\simeq$\,20\,--\,100\,\kms and pressure of $P/{\rm K}_B\simeq 10^{7.6}\,\rm K\,$\cc (i.e., $> 1000 \times$ with respect to the Milky Way), similar to those found in nearby and $z$\ssim2 gas-rich starburst galaxies.
%
In addition, we perform stability analysis to understand the origin and dynamical properties of MCCs. We find that MCCs are globally stable in the main disk of \flower. Densest regions where \SF is expected to take place in clumps and cores on even smaller scales instead  have lower $\alpha_{\rm vir}$ and Toomre-$Q$ values.
%
Detailed studies of the star-forming gas dynamics at EoR thus requires a spatial resolution of $\lesssim$40\,pc ($\simeq$\,0$\farcs$01), which is within reach with the Atacama Large (sub-)Millimeter Array (ALMA) and the Next Generation Very Large Array (ngVLA).
\end{abstract}
\keywords{methods: data analysis --
          galaxies: high-redshift --
          galaxies: ISM --
          galaxies: evolution --
          galaxies: formation --
          galaxies: starburst --
          stars: formation}

@@

To understand the physical properties of MCCs in early galaxies, we have undertaken a detailed numerical study whose aim is to characterize the dynamical properties of the star-forming MCCs in prototypical (i.e., $L^*$) galaxies in the Epoch of Reionization (EoR).

@@

% address the biggest MCC = molecular disk of \flower
%mm2 [This and following figures would be better with the  color bar
%from Figure 7 rather than the prominent stars at the bottom...]
\begin{figure}
\centering
\includegraphics[width=0.5\textwidth]{\figpath/cloud-mass_fgas.pdf}
\caption{Cloud mass and stellar-to-gas mass ratio of MCCs identified in the accreting phase of \flower with different $n_{\rm cut}$. The most massive structures identified with a low $f_{\rm gas}$ correspond to the molecular disk/arms of \flower (see also \Fig{vv}).
%
Thus, these structures encompass a large amount of stellar mass that is already assembled in \flower. These disk/arm structures are excluded in the discussion of MCC dynamics in the remainder of this paper.
\label{fig:stellarRatio16}}
\end{figure}

@@

This also implies that \obs on spatially resolution better than $\simeq$40\,pc are needed to examine the truly star-forming structures (cores), and thus, \SF in the first galaxies. Such resolution is in principle within reach using ALMA, for which 0\farcs01 resolution images can be attained using the high frequency bands. With the Next Generation VLA (ngVLA\footnote{\url{http://library.nrao.edu/public/memos/ngvla/NGVLA\_21.pdf}}), which is approximately ten times more sensitive than ALMA, direct observations of molecular gas clouds in galaxies at the EoR on cloud scales will be possible via the higher-$J$ rotational transitions of CO in the 2030s.
% 0.005" highest freq
In the meantime, cosmological zoom-in simulations, such as \ncode{Serra}, while inherently limited in galaxy statistics and are subject to the sub-grid models adopted, serve as a useful tool for examining and making predictions on the morphology and dynamics of the molecular ISM of the first galaxies.

%-------------------------------------------------------
%                                Back matters
%-------------------------------------------------------
% ACKNOWLEDGEMENTS
%-------------------------------------
% \acknowledgements

\section*{Acknowledgements}
We thank Jens Kauffmann, Thushara Pillai, and Mark Swinbank for sharing their data, and Caitlin Casey for useful discussions.
%
TKDL gratefully acknowledges support by the NSF through award SOSPA4-009 from the NRAO and support from the Simons Foundation.
%
AF acknowledges support from the ERC Advanced Grant INTERSTELLAR
H2020/740120.
% mm
M-MML acknowledges partial support by the NSF through award
AST18-15461, and hospitality from the Insitute for Theoretical
Astrophysics at the University of Heidleberg.
%
This work is based on a project developed at the Kavli Summer Program in Astrophysics (KSPA) held at the Center for Computational Astrophysics of the Flatiron Institute in 2018. The program was co-funded by the Kavli Foundation and the Simons Foundation.
%
We thank the KSPA Scientific and Local Organizing Committees, and the program founder, Pascale Garaud, for supporting the genesis of this work.
%
We also thank the Center for Cosmology and Particle Physics at the New York University for their hospitality in hosting us after the steam pipe explosion in NYC during the KSPA.
%
This research has made use of NASA's Astrophysics Data System Bibliographic Services.
%
We acknowledge use of the Python programming language \citep{VanRossum1991}, Astropy \citep{astropy}, Cython \citep{behnel2010cython}, Matplotlib \citep{Hunter2007}, NumPy \citep{VanDerWalt2011}, \ncode{Pymses} \citep{Labadens2012}, SciPy \citep{scipyref}, and \ncode{yt} \citep{Smith09a,Turk11a}.
%
%and \ncode{yt}, a community-developed Python package for analyzing simulation data in Astronomy.

>>>>

\begin{abstract}
We study the properties of molecular cloud complexes (MCCs) at the epoch of reionization using cosmological zoom-in simulations. We identify MCCs in a $z\simeq 6$ prototypical galaxy (``\flower'') using an H$_2$ density-based clump finder.
%
We compare their mass, size, velocity dispersion, gas surface density, and virial parameter ($\alpha_{\rm vir}$) to \obs. In \flower, the typical MCC mass and size are $M_{\rm gas} \simeq 10^{6.5}$\,\Msun and $R\simeq45-100$\,pc, which are comparable to those found in nearby spirals and starburst galaxies. MCCs are highly supersonic and supported by turbulence, with r.m.s velocity dispersions of $\sigma_{\rm gas}\simeq$\,20\,--\,100\,\kms and pressure of $P/{\rm K}_B\simeq 10^{7.6}\,\rm K\,$\cc (i.e., $> 1000 \times$ with respect to the Milky Way), similar to those found in nearby and $z$\ssim2 gas-rich starburst galaxies.
%
In addition, we perform stability analysis to understand the origin and dynamical properties of MCCs. We find that MCCs are globally stable in the main disk of \flower. Densest regions where \SF is expected to take place in clumps and cores on even smaller scales instead  have lower $\alpha_{\rm vir}$ and Toomre-$Q$ values.
%
Detailed studies of the star-forming gas dynamics at the epoch of reionization thus require a spatial resolution of $\lesssim$40\,pc ($\simeq$\,0$\farcs$01), which is within reach with the Atacama Large (sub-)Millimeter Array %mm2 (ALMA)
and the Next Generation Very Large Array.
%mm2 (ngVLA).
\end{abstract}
\keywords{methods: data analysis --
          galaxies: high-redshift --
          galaxies: ISM --
          galaxies: evolution --
          galaxies: formation --
          galaxies: starburst --
          stars: formation}

@@

To understand the physical properties of MCCs in early galaxies, we have undertaken a detailed numerical study whose aim is to characterize the dynamical properties of the star-forming MCCs in prototypical (i.e., $L^*$) galaxies in the epoch of reionization (EoR).

@@

\begin{figure}
\centering
\includegraphics[width=0.5\textwidth]{\figpath/cloud-mass_fgas.pdf}
\caption{Cloud mass and gas mass ratio of MCCs identified in the accreting phase of \flower using different $n_{\rm cut}$ (see colorbar). The most massive structures identified with a low $f_{\rm gas}$ correspond to the molecular disk/arms of \flower (see also \Fig{vv}).
%
Thus, these structures encompass a large amount of stellar mass that is already assembled in \flower. These disk/arm structures are excluded in the discussion of MCC dynamics in the remainder of this paper.
\label{fig:stellarRatio16}}
\end{figure}

@@

This also implies that \obs with spatial resolution better than $\simeq$40\,pc are needed to examine the truly star-forming structures (cores), and thus, \SF in the first galaxies. Such resolution is in principle within reach using ALMA, for which 0\farcs01 resolution images can be attained using the high frequency bands. With the Next Generation VLA%
%mm2 (ngVLA
\footnote{\url{http://library.nrao.edu/public/memos/ngvla/NGVLA\_21.pdf}}, which is approximately ten times more sensitive than ALMA, direct observations of molecular gas clouds in galaxies at the EoR on cloud scales will be possible via the higher-$J$ rotational transitions of CO in the 2030s.
% 0.005" highest freq
In the meantime, cosmological zoom-in simulations, such as \ncode{Serra}, while inherently limited in galaxy statistics, and
% mm2 subject to
    dependent on
the sub-grid models adopted, serve as a useful tool for examining and making predictions on the morphology and dynamics of the molecular ISM of the first galaxies.

%                                Back matters
%-------------------------------------------------------
% ACKNOWLEDGEMENTS
%-------------------------------------
% \acknowledgements

\section*{Acknowledgements}
We thank Jens Kauffmann, Thushara Pillai, and Mark Swinbank for sharing their data, and Caitlin Casey for useful discussions.
%
TKDL gratefully acknowledges support by the NSF through award SOSPA4-009 from the NRAO and support from the Simons Foundation.
%
AF acknowledges support from the ERC Advanced Grant INTERSTELLAR
H2020/740120.
% mm
M-MML acknowledges partial support by the NSF through award
AST18-15461, and hospitality from the Insitute for Theoretical
Astrophysics at the University of Heidleberg.
%
This work is based on a project developed at the Kavli Summer Program in Astrophysics (KSPA) held at the Center for Computational Astrophysics of the Flatiron Institute in 2018. The program was co-funded by the Kavli Foundation and the Simons Foundation.
%
We thank the KSPA Scientific and Local Organizing Committees, and the program founder, Pascale Garaud, for supporting the genesis of this work.
%
We also thank the Center for Cosmology and Particle Physics at the New York University for their hospitality in hosting us after the steam pipe explosion in NYC during the KSPA.
%
This research has made use of NASA's Astrophysics Data System Bibliographic Services.
%
% mm2 We acknowledge use of the

\software{Python \citep{VanRossum1991}, Astropy \citep{astropy}, Cython \citep{behnel2010cython}, Matplotlib \citep{Hunter2007}, NumPy \citep{VanDerWalt2011}, \ncode{PyMSES} \citep{Labadens2012}, SciPy \citep{scipyref}, and \ncode{yt} \citep{Smith09a,Turk11a}}
%
%and \ncode{yt}, a community-developed Python package for analyzing simulation data in Astronomy.
